Intentional release of native species undermines ecological stability

### 2022-12-24 rmd/manuscript.Rmd
edit: manuscript

The intentional release, however, can have far-reaching consequences than previously thought, as the impacts can propagate through a diversity of ecological interactions [@careyNativeInvadersChallenges2012]. This recognition has stimulated a discussion of how the massive introduction of native species alters short-term ecosystem dynamics [@hasegawaAreHatcheryChum2018; @nelsonEcologicalImplicationsChanging2019]. Yet, current debates overlook the fact that we lack a quantitative assessment of the community-wide impact in the long-term. Ecological theory suggests that the stable coexistence of competing species through density-dependent feedback underpins the emergent stability of ecological communities [@tilmanEcologicalConsequencesChanges1999; @carrollNicheFitnessDifferences2011a; @loreauDoesFunctionalRedundancy2004]. Intentional release may disrupt the sensitive balance of species interactions because it introduces unnaturally high numbers of individuals into the wild [@laikreCompromisingGeneticDiversity2010; @careyNativeInvadersChallenges2012; @kitadaEconomicEcologicalGenetic2018]. Hence, this form of species management may intervene in the ecological process that allows competing species to coexist, ultimately degrading the long-term community stability. Evidence for this hypothesis is lacking, however.

@@

The intentional release had little or a stabilizing influence on community dynamics when carrying capacity was sufficiently large ($K=$ `r max(list_param[[2]]$k)`; **Figures S7 -- S10**). In particular, intentional release increased the enhanced species with a low population growth rate ($r_1=$ `r min(list_param[[2]]$r1)`; **Figure S7**), suggesting that weak intraspecific competition within the enhanced species (= $r_1 K^{-1}$) is needed for intentional release to be effective. The weak-competition scenario is likely in reintroduction programs that aim to replenish populations of extirpated species or long-lived endangered species. In fact, some of the best evidence for successful intentional release comes from reintroduction efforts for such species [@deguchiTranslocationHandrearingResult2017; @jonesRestorationMauritiusKestrel1995; @okahisaEvaluationOngoingCrested2022; @yackulicAssessingPopulationImpacts2021]. In the Sado island (Japan), for example, the once-extirpated Crested Ibis *Nipponia nippon* showed exponential growth of the population since the initial reintroduction of ten captive-bred individuals in 2008 [@okahisaEvaluationOngoingCrested2022]. Similarly, translocations of Humpback chub, a federally-listed fish species in the US, seem highly cost-effective in enhancing their tributary populations in the Little Colorado River [@yackulicAssessingPopulationImpacts2021]. However, the preponderance of unsuccessful reintroduction programs implies that the theoretical requirements for successful intentional release are not always met [@fischerAssessmentPublishedResults2000; @haagNorthAmericanFreshwater2012; @strayerEssayMakingMost2019]. Indeed, non-negligible numbers of projects seem to release captive-bred individuals into unsuitable habitats with compromised carrying capacities [@fischerAssessmentPublishedResults2000; @haagNorthAmericanFreshwater2012]. Our analysis suggests that such programs could rather impair biodiversity without noticeable demographic benefit to the enhanced species. Pre-release examination and restoration of environmental capacity may be key to successful release programs with minimal impacts on other community members.

@@

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of the intentional release of masu salmon (**Figure** 1)**E**) on the long-term stability of stream fish communities in Hokkaido, Japan. In the protected watersheds (all separated by the ocean; **Figure 1C, D**), a long-term program exists to monitor stream fish communities along with the official release records. The release of masu salmon began in the 1950s, with the annual average ranging from `r round(min(df_stock_mu$mean_stock), 2)` (no release) to `r round(max(df_stock_mu$mean_stock), 2)` million fish across watersheds during 1999 -- 2019. The majority of release occurs in spring near fish survey locations, after which salmon fry stays in freshwater for at least one year with other stream fishes (see **Figure S16** for a co-occurrence matrix in our dataset). Several of the watersheds are not subject to intentional release aiming to preserve regional populations of masu salmon, while the release numbers at other watersheds vary due mainly to logistical reasons. Therefore, the study system sets the stage for a "natural experiment" to test our theoretical predictions. We used the data from 1999 to 2019 at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` independent watersheds (see **Methods** for selection criteria). Using hierarchical Bayesian models, we quantified the effect of intentional release on community dynamics while accounting for the potential effects of climates and local abiotic factors.

@@

To gain better insights into the driving mechanisms, we fitted a multi-species Ricker model to a subset of our time-series data (nine sites with few missing observations and no intentional release; **Methods**). The parameter estimates of this model are directly comparable to those used in the theoretical Ricker model. The estimates of interspecific competition coefficients $\alpha_{ij}$ ranged `r op(min(df_pd$alpha_prime), 2)` -- `r op(max(df_pd$alpha_prime), 2)` with the median of `r op(median(df_pd$alpha_prime), 2)` (**Figure** 3), suggesting that competition is a likely mechanism as our theory assumes (see **Figure S17** for all pairwise estimates). Our statistical inference is supported by previous studies. For example, competition between masu salmon and other salmonid species is well documented, and its form can be interference or exploitative [@miyasakaFoodHabitDivergence2003; @hasegawaComparisonCompetitiveAbility2004; @moritaImpactExoticTrout2004]. Diets of salmonid species are innately overlapped, and they interfere with each other to claim profitable foraging spots for drift insects [@hasegawaComparisonCompetitiveAbility2004] or cruise over a large area in search of prey [@hasegawaEffectFlowRegime2010]. Exploitation competition between masu salmon and benthic fish (e.g., sculpins and stone loaches in our study system) is also likely because subordinate masu salmon in the dominance hierarchy consumes benthic macroinvertebrates [@satoStagespecificFunctionalResponses2014] that many benthic fish species prey on [@natsumedaEvaluationEcologicalFeatures2010; @natsumedaDietFoodSelectivity2012]. Importantly, hatchery masu salmon may intensify intra- and interspecific competition in the wild [@hasegawaWildMasuSalmon2018; @sahashiEvaluationContributionHatchery2015; @hasegawaPulsedSuppliesSmall2022]. Hatchery salmon are larger and more aggressive than wild individuals of a comparable life stage, increasing the likelihood of intense intra- and interspecific competition [@weberInteractionsHatcheryWild2003]. Indeed, field and experimental studies confirmed that hatchery masu salmon competed with wild masu salmon and other stream fishes during their freshwater life stage [@hasegawaWildMasuSalmon2018; @sahashiEvaluationContributionHatchery2015; @hasegawaPulsedSuppliesSmall2022].

@@

Human activities are strictly regulated in the protected watersheds, and the near-intact environment (**Figure S10**) may have helped uncover the qualitative agreement between theory and empirical patterns. For example, any form of exploitation, including angling, is prohibited in these areas. In addition, exclusive permission from the governer of Hokkaido is required to physically alter in-stream and riparian habitats. Therefore, unmeasured human influences should have minimal influences on the observed relationships. Further, we have statistically controlled natural variation in environmental factors (see **Methods**). Nevertheless, our results must be viewed with caution: we cannot exclude the possibility of spurious correlations due to the inherent nature of correlative analysis and field research. Since an experimental approach is nearly impossible at this spatial scale, the application of novel causal inferences (e.g., convergent cross mapping) [@sugiharaDetectingCausalityComplex2012; @ushioFluctuatingInteractionNetwork2018; @suzukiDecomposingPredictabilityIdentify2022] may be a potential alternative to confirm the causal relationship between intentional release and community dynamics. Unfortunately, these methods were not applicable to our data because they require a longer and non-sparse time series to yield robust results.

@@

**Fish release.** The release of masu salmon began in the 1950s. The duration of captivity ranged from zero (wild origin) to seven generations, and the majority is released at two to four generations in captivity. Although the program aims to release a constant number of hatchery fish for each watershed, uncontrollable factors (e.g., disease, budget allocation) produce some variation in the number of releases among years. Hatchery fish are released in spring (fry and smolt stages) and fall (juvenile stage). Most fish are released near the fish survey sites. However, the number of fish released at each site was reported as an aggregate for each watershed, and the site-specific information of release is unavailable. For each release stage, we assembled annual records of intentional release (the number of fish released; 1999 -- 2019) from annual investigations by the Japan Fisheries Research and Education Agency and Salmon and Freshwater Fisheries Research Institute. During the study period, the majority of release took place in spring at a fry stage (fry : juvenile : smolt = `r pull(df_stock_stage, ratio)[1]` : `r op(pull(df_stock_stage, ratio)[2])` : `r op(pull(df_stock_stage, ratio)[3])`)

@@

We used linear regression to quantify the impact of the intentional release on community dynamics. Although our focus is intentional release, each model included climatic and local abiotic variables to account for important environmental differences among sites. Specifically, we developed the following linear regression model taking either species richness (the number of species present during the observation period), mean, or SD as a response variable $y_s$.

@@

**Model fitting.** We fitted the models to the data using JAGS version 4.1.0 through *runjags* package version 2.2.0-2 in R [@denwoodRunjagsPackageProviding2016]. We assigned vague or weakly informative priors to parameters (**Table S**3). Four Markov chain Monte Carlo (MCMC) chains were run until parameter estimates converged. The total MCMC iterations ranged from `r min(df_mcmc$n_total_mcmc)` to `r max(df_mcmc$n_total_mcmc)`, in which MCMC samples were saved every `r unique(df_mcmc$n_thin)` steps for the calculation of posterior distributions after the initial `r max(df_mcmc$n_burn)` burn-in period. Convergence was assessed by examining whether the $\hat{R}$ indicator of each parameter approached $<1.1$ [@gelmanDataAnalysisUsing2007]. Data manipulation and analysis were performed in R version 4.2.1 [@rcoreteamLanguageEnvironmentStatistical2021]. Parameter estimates were summarized in **Tables S**4 -- **S**8.

>>>>

The intentional release, however, can have far-reaching consequences than previously thought, as the impacts can propagate through a diversity of ecological interactions [@careyNativeInvadersChallenges2012]. This recognition has stimulated a discussion of how the massive introduction of native species alters short-term ecosystem dynamics [@hasegawaAreHatcheryChum2018; @nelsonEcologicalImplicationsChanging2019; @cucheroussetAreDomesticatedFreshwater2020]. Yet, current debates overlook the fact that we lack a quantitative assessment of the community-wide impact in the long-term. Ecological theory suggests that the stable coexistence of competing species through density-dependent feedback underpins the emergent stability of ecological communities [@tilmanEcologicalConsequencesChanges1999; @carrollNicheFitnessDifferences2011a; @loreauDoesFunctionalRedundancy2004]. Intentional release may disrupt the sensitive balance of species interactions because it introduces unnaturally high numbers of individuals into the wild [@laikreCompromisingGeneticDiversity2010; @careyNativeInvadersChallenges2012; @kitadaEconomicEcologicalGenetic2018]. Hence, this form of species management may intervene in the ecological process that allows competing species to coexist, ultimately degrading the long-term community stability. Evidence for this hypothesis is lacking, however.

@@

The intentional release had little or a stabilizing influence on community dynamics when carrying capacity was sufficiently large ($K=$ `r max(list_param[[2]]$k)`; **Figures S7 -- S10**). In particular, intentional release increased the enhanced species with a low population growth rate ($r_1=$ `r min(list_param[[2]]$r1)`; **Figure S7**), suggesting that weak intraspecific competition within the enhanced species (= $r_1 K^{-1}$) is needed for intentional release to be effective. The weak-competition scenario is likely in reintroduction of extirpated species or long-lived endangered species. In fact, some of the best evidence for successful intentional release comes from conservation programs of such species [@deguchiTranslocationHandrearingResult2017; @jonesRestorationMauritiusKestrel1995; @okahisaEvaluationOngoingCrested2022; @yackulicAssessingPopulationImpacts2021]. In the Sado island (Japan), for example, the once-extirpated Crested Ibis *Nipponia nippon* showed exponential growth of the population since the initial reintroduction of ten captive-bred individuals in 2008 [@okahisaEvaluationOngoingCrested2022]. Similarly, translocations of Humpback chub, a federally-listed fish species in the US, seem highly cost-effective in enhancing their tributary populations in the Little Colorado River [@yackulicAssessingPopulationImpacts2021]. However, the preponderance of unsuccessful reintroduction in conservation implies that the theoretical requirements for successful intentional release are not always met [@fischerAssessmentPublishedResults2000; @haagNorthAmericanFreshwater2012; @strayerEssayMakingMost2019]. Indeed, non-negligible numbers of projects seem to release captive-bred individuals into unsuitable habitats with compromised carrying capacities [@fischerAssessmentPublishedResults2000; @haagNorthAmericanFreshwater2012]. Our analysis suggests that such programs could rather impair biodiversity without noticeable demographic benefit to the enhanced species. Pre-release examination and restoration of environmental capacity may be key to successful release programs with minimal impacts on other community members.

@@

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of the intentional release of masu salmon (**Figure** 1**E**) on the long-term stability of stream fish communities in Hokkaido, Japan. In the protected watersheds (all separated by the ocean; **Figure 1C, D**), a long-term program exists to monitor stream fish communities along with the official release records. The release of masu salmon began in the 1950s; the release numbers at other watersheds vary due mainly to logistical reasons, while several of the watersheds are not subject to intentional release to preserve regional populations. This resulted in the average annual release ranging from `r round(min(df_stock_mu$mean_stock), 2)` (no release) to `r round(max(df_stock_mu$mean_stock), 2)` million fish across watersheds during 1999 -- 2019. The majority of release occurs in spring near fish survey locations, after which salmon fry stays in freshwater for at least one year with other stream fishes (see **Figure S16** for a co-occurrence matrix in our dataset). Therefore, the study system sets the stage for a "natural experiment" to test our theoretical predictions. We used the data from 1999 to 2019 at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` independent watersheds (see **Methods** for selection criteria). Using hierarchical Bayesian models, we quantified the effect of intentional release on community dynamics while accounting for the potential effects of climates and local abiotic factors.

@@

To gain better insights into the driving mechanisms, we fitted a multi-species Ricker model to a subset of our time-series data (nine sites with few missing observations and no intentional release; **Methods**). The parameter estimates of this model are directly comparable to those used in the theoretical Ricker model. The estimates of interspecific competition coefficients $\alpha_{ij}$ ranged `r op(min(df_pd$alpha_prime), 2)` -- `r op(max(df_pd$alpha_prime), 2)` with the median of `r op(median(df_pd$alpha_prime), 2)` (**Figure** 3), suggesting that competition is a likely mechanism as our theory assumes (see **Figure S17** for all pairwise estimates). Our statistical inference is supported by previous studies. For example, competition between masu salmon and other salmonid species is well documented, and its form can be interference or exploitative [@miyasakaFoodHabitDivergence2003; @hasegawaComparisonCompetitiveAbility2004; @moritaImpactExoticTrout2004]. Diets of salmonid species are innately overlapped, and they interfere with each other to claim profitable foraging spots for drift insects [@hasegawaComparisonCompetitiveAbility2004] or cruise over a large area in search of prey [@hasegawaEffectFlowRegime2010]. Exploitation competition between masu salmon and benthic fish (e.g., sculpins and stone loach in our study system) is also likely because subordinate masu salmon in the dominance hierarchy consumes benthic macroinvertebrates [@satoStagespecificFunctionalResponses2014] that many benthic fish species prey on [@natsumedaEvaluationEcologicalFeatures2010; @natsumedaDietFoodSelectivity2012]. Importantly, hatchery masu salmon may intensify intra- and interspecific competition in the wild [@hasegawaWildMasuSalmon2018; @sahashiEvaluationContributionHatchery2015; @hasegawaPulsedSuppliesSmall2022]. Hatchery salmon are larger and more aggressive than wild individuals of a comparable life stage, increasing the likelihood of intense intra- and interspecific competition [@weberInteractionsHatcheryWild2003]. Indeed, field and experimental studies confirmed that hatchery masu salmon competed with wild masu salmon and other stream fishes during their freshwater life stage [@hasegawaWildMasuSalmon2018; @sahashiEvaluationContributionHatchery2015; @hasegawaPulsedSuppliesSmall2022].

@@

Two unique strengths in our dataset may have helped uncover the qualitative agreement between theory and empirical patterns. First, human activities are strictly regulated in the protected watersheds (**Figure S10**). Any form of exploitation, including angling, is prohibited in these areas. In addition, exclusive permission from the governer of Hokkaido is required to physically alter in-stream and riparian habitats. Therefore, unmeasured human influences should have minimal influences on the observed relationships. As such, we were able to confine potential confounding factors to natural environmental variation, which were properly accounted in our statistical model (see **Methods**). Second, the sampling method in this monitoring program (a combination of cast net and electrofishing) is carefully chosen to capture stream fishes with different microhabitat preference (e.g., water-column vs. bottom). Thus, our data is robust to potential bias caused by sampling methods. Nevertheless, our results must be viewed with some caution: we cannot exclude the possibility of spurious correlations due to the inherent nature of correlative analysis and field research. Since an experimental approach is nearly impossible at this spatial scale, the application of novel causal inferences (e.g., convergent cross mapping) [@sugiharaDetectingCausalityComplex2012; @ushioFluctuatingInteractionNetwork2018; @suzukiDecomposingPredictabilityIdentify2022] may be a potential alternative to confirm the causal relationship between intentional release and community dynamics. Unfortunately, these methods were not applicable to our data because they require a longer and non-sparse time series to yield robust results.

@@

**Fish release.** The release of masu salmon began in the 1950s. The duration of captivity ranged from zero (wild origin) to seven generations, and the majority is released at two to four generations in captivity. Although the program aims to release a constant number of hatchery fish for each watershed, uncontrollable factors (e.g., disease, budget allocation) produce some variation in the number of releases among years. Hatchery fish are released in spring (fry and smolt stages) and fall (juvenile stage). Most fish are released near the fish survey sites. However, the number of fish released was reported as an aggregate for each watershed, and the site-specific information of release is unavailable. For each release stage, we assembled annual records of intentional release (the number of fish released; 1999 -- 2019) from annual investigations by the Japan Fisheries Research and Education Agency and Salmon and Freshwater Fisheries Research Institute. During the study period, the majority of release took place in spring at a fry stage (fry : juvenile : smolt = `r pull(df_stock_stage, ratio)[1]` : `r op(pull(df_stock_stage, ratio)[2])` : `r op(pull(df_stock_stage, ratio)[3])`)

@@

We used linear regression to quantify the impact of the intentional release on community dynamics. Although our focus is intentional release, each model included climatic and local abiotic variables to account for important environmental differences among sites. Specifically, we developed the following linear regression model taking either taxonomic richness (the number of taxa present during the observation period), mean, or SD as a response variable $y_s$.

@@

**Model fitting.** We fitted the models to the data using JAGS version 4.1.0 through *runjags* package version 2.2.0-2 in R [@denwoodRunjagsPackageProviding2016]. We assigned vague or weakly informative priors to parameters (**Table S**3). Four Markov chain Monte Carlo (MCMC) chains were run until parameter estimates converged. The total MCMC iterations ranged from `r min(df_mcmc$n_total_mcmc) %>% as.character()` to `r max(df_mcmc$n_total_mcmc) %>% as.character()`, in which MCMC samples were saved every `r unique(df_mcmc$n_thin)` steps for the calculation of posterior distributions after the initial `r max(df_mcmc$n_burn)` burn-in period. Convergence was assessed by examining whether the $\hat{R}$ indicator of each parameter approached $<1.1$ [@gelmanDataAnalysisUsing2007]. Data manipulation and analysis were performed in R version 4.2.1 [@rcoreteamLanguageEnvironmentStatistical2021]. Parameter estimates were summarized in **Tables S**4 -- **S**8.
